---
title: "Penguins EDA"
format:
  html:
    code-fold: true
---

## Penguin Size and Mass by Sex and Species

```{R}
library(caret)
```

```{r}
library(palmerpenguins)
library(dplyr)
library(ggplot2)

df <- palmerpenguins::penguins
```

```{r}
head(df)
```

```{r}
df %>%
  group_by(species, sex) %>%
  summarise(
    across(
      where(is.numeric), 
      \(x) mean(x, na.rm = TRUE)
      )
    ) %>%
  knitr::kable()
```

## Penguin Size vs Mass by Species

```{r}
df %>%
  ggplot(aes(x = bill_length_mm, y = body_mass_g, color = species)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

## That's very pretty. I'd like to know some other things though, and I like python more

```{python}
df.info()
```

## That's a clean dataset! Let's see what the species and island breakdowns are.

```{python}
print(pd.DataFrame(df['species'].value_counts()))
```

```{python}
print(pd.DataFrame(df['island'].value_counts()))
```

```{R}
ggplot(data = penguins, aes(x = island, fill = species)) + 
  geom_bar(position = "stack") +
  labs(title = "Count of Penguin Species by Island",
       x = "Island",
       y = "Count") +
  theme_minimal()
```

# Neato! Now let's see the same breakdown as before, but for island instead of population

```{R}
df %>%
  ggplot(aes(x = bill_length_mm, y = body_mass_g, color = island)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

# Interesting. Seems less pronounced than species.Also, the groupings that appear are probably just masked versions of the species effects.
